---
title: "Reproducible Research: Peer Assessment 1"
author: "Sergey Nazarov"
date: "Wednesday, August 12, 2015"
output: html_document
---

## Loading and preprocessing the data

```{r}
unzip("./repdata-data-activity.zip")
data <- read.csv("./activity.csv")
```

looking at data
```{r}
summary(data)
head(data)
str(data)
```

## What is mean total number of steps taken per day?

here i aggregate step numbers by date, which i will use later in histogram
```{r}
aggregate(steps ~ date, data = data, sum)
```

making histogram steps by day. at first look it looks like normal distribution just
with more thick first half
```{r histogram}
hist(aggregate(steps ~ date, data = data, sum)$steps, main = "Histogram of steps by day",
     xlab = "number of steps")
```

here we can look at histogram more closely, using "breaks", and we see something
interseting - days with extra activity (more than 20000 steps)
```{r}
hist(aggregate(steps ~ date, data = data, sum)$steps, breaks = 10, main = "Histogram of steps by day",
     xlab = "number of steps")
```

### Mean number of steps by day
```{r}
mean(aggregate(steps ~ date, data = data, sum)$steps)
```

### Median number of steps by day
```{r}
median(aggregate(steps ~ date, data = data, sum)$steps)
```


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?







### Version information
```{r}
sessionInfo()
```