---
title: "Reproducible Research: Peer Assessment 1"
author: "Sergey Nazarov"
date: "Wednesday, August 12, 2015"
output: html_document
---

## Loading and preprocessing the data

```{r}
unzip("./repdata-data-activity.zip")
data <- read.csv("./activity.csv")
```

looking at data
```{r, results='hide'}
summary(data)
head(data)
str(data)
```


## What is mean total number of steps taken per day?

here i aggregate step numbers by date, which i will use later in histogram
```{r, results='hide'}
aggregate(steps ~ date, data = data, sum)
```

we can look at histogram, using "breaks" = 10, and we see something
interseting - some days with extra activity (more than 20000 steps)
```{r}
hist(aggregate(steps ~ date, data = data, sum)$steps, breaks = 10, main = "Histogram of steps by day",
     xlab = "number of steps")
```

### Mean number of steps by day
```{r}
mean(aggregate(steps ~ date, data = data, sum)$steps)
```

### Median number of steps by day
```{r}
median(aggregate(steps ~ date, data = data, sum)$steps)
```

## What is the average daily activity pattern?

```{r}
dailypattern <- aggregate(steps ~ interval, data = data, mean)
plot(dailypattern, type = "l")
```

finding maximum number of steps
```{r}
dailypattern[dailypattern$steps == max(dailypattern$steps), ]
```
the maximum number of steps contains interval # 835 (on average across all the days in the dataset)
it's actually 8:35 AM - jogging or hurrying to work? :)

## Imputing missing values
lets see how many missing values are in the dataset
```{r}
sum(is.na(data$steps))
sum(is.na(data$steps))/nrow(data)
```
near 13% is missing

for this task i create a separate data frame *"data without NA"* in which i fill NA values
using the mean for the correspondent 5-minute interval
```{r}
datawithoutNA <- data

for (i in 1:nrow(datawithoutNA)) {
        if (is.na(datawithoutNA$steps[i]) == TRUE) {
                datawithoutNA$steps[i] <- dailypattern$steps[dailypattern$interval == datawithoutNA$interval[i]]
        }
}
```
just checking there is no NA's
```{r}
sum(is.na(datawithoutNA$steps)) 
```

here i aggregate step numbers by date, which i will use later in histogram2
```{r, results='hide'}
aggregate(steps ~ date, data = datawithoutNA, sum)
```

here we can look at histogram
```{r}
hist(aggregate(steps ~ date, data = datawithoutNA, sum)$steps, breaks = 10, main = "Histogram of steps by day",
     xlab = "number of steps")
```

### Mean number of steps by day without NA
```{r}
mean(aggregate(steps ~ date, data = datawithoutNA, sum)$steps)
```

### Median number of steps by day without NA
```{r}
median(aggregate(steps ~ date, data = datawithoutNA, sum)$steps)
```
### the is no difference between mean number of steps with and without NA


## Are there differences in activity patterns between weekdays and weekends?
first I create a new dataset using data without missing values and transform date from factor to time 
```{r}
dataweek <- datawithoutNA
dataweek$date <- strptime(dataweek$date, "%Y-%m-%d")
```
then I add a new variable using weekdays() function
```{r}
Sys.setlocale("LC_TIME", "English")
dataweek$day <- weekdays(dataweek$date, abbreviate = FALSE)
```
then I transform weekay names into two variables: weekend and weekday 
```{r}
for (i in 1:nrow(dataweek)) {
        if (dataweek$day[i] == "Saturday" | dataweek$day[i] == "Sunday") {
                dataweek$day[i] <- "weekend"
        } else {dataweek$day[i] <- "weekday"}
}
```
Here I create two separate datasets (weekdays and weekend) with patterns on average 
across all the days in each dataset
```{r}
weekendpattern <- aggregate(steps ~ interval, data = dataweek[dataweek$day == "weekend",], mean)
weekdaypattern <- aggregate(steps ~ interval, data = dataweek[dataweek$day == "weekday",], mean)
```
making plot for each dataset
```{r}
par(mfrow=c(2 , 1))
par(mar = c(2, 2, 2, 1))
with(weekdaypattern,  {plot(interval, steps, type = "l", 
                        xlab = "", ylab = "", main = "weekday")})
with(weekendpattern,  {plot(interval, steps, type = "l", 
                            xlab = "", ylab = "", main = "weekend")})
```
  
we can see difference in daily activity pattern between weekdays and weekend  
- later start of activity on weekends (presumably late awakening)  
- lower level of active during the day on weekdays then weekends. may be becouse of sitting type of job
        of the subject



### Version information
```{r}
sessionInfo()
```